Practical Parametricity for GADTs

### 2020-09-03 20-gadts/inf-rules.tex
added more intution about par for GADTs at start of main paper

{\color{red} We should ask what, intuitively, it means to a GADT to be
  ``parametric'', i.e., uniform in its type indices. For ADTs and
  nested types it means that the datatype elements are constructed in
  exactly the same ways at each type. For GADTs, it means that as
  well, but here we have to take into account the inhabited instances
  as well. [Of course we did not have to do that ADTs and nested types
    because {\em all} instances of ADTs and nested types are
    inhabited.] So, intuitively, we consider a GADT to be
  ``parametric'' if we can construct the inhabitants of all of the
  inhabited type instances in a type-independent way. We capture this
  in our calculus by requiring that $\emptyset;\ol\beta \vdash K$ (and
  that $K$ does not involve any constants?) in $\Gamma;\Phi \vdash
  (\Lan^{\ol{\alpha^0}}_{\ol K} F)\, \ol A$. Then to get that the real
  live Lans interpreting $\Lan$-constructs are $\omega$-cocontinuous
  we also require that $K$ is a polynomial functor in the $\alpha$s
  that does not involve any constant functors. This means that our
  constructors have types of the form $F\ol\beta \to F(K\ol\beta)$
  with no constants in $K$, so that exactly the corresponding
  instances of the GADT are inhabited for every instantiation of the
  $\alpha$s, and the objects for each instantiation of the $\alpha$s
  are constructed in exactly the same way. That is, the inhabited
  instances and the construction of the data in those instances are
  type independent.}

>>>>

{\color{red} We should ask what, intuitively, it means to have
  parametricity for a calculus with GADTs. In particular it should
  mean that any polymorphic function that takes a GADT as input should
  operate uniformly over all type instances of the GADT. That means
  that the GADT must itself be ``parametric'', i.e., uniform in its
  type indices. But what does this mean? For ADTs and nested types it
  means that the datatype elements are constructed in exactly the same
  ways at each type instance. For GADTs, it means that as well, but
  here we also have to take into account which instances are actually
  inhabited, since not all type instances of a GADT are inhabited.
  [Of course we did not have to do this for ADTs and nested types
    because {\em all} instances of ADTs and nested types are
    inhabited.] So, intuitively, we consider a GADT to be
  ``parametric'' if 1) the corresponding instances of the GADT are
  inhabited for for every instantiation of its type parameters
  $\ol\alpha$, and 2) the objects for each instantiation of the
  $\alpha$s are constructed in exactly the same way. That is, which
  instances are inhabited, as well as how the data in each of those
  instances are constructed, is type independent. We capture this in
  our calculus by requiring that, for $\Gamma;\Phi \vdash
  (\Lan^{\ol{\alpha^0}}_{\ol K} F)\, \ol A$, we have that
  $\emptyset;\ol\beta \vdash K$, and that $K$ does not involve any
  constants. (The latter ensures that a GADT with a constructor of
  type $\alpha \to Int$ is not parametric, which makes sense since
  there will be more ways to construct Int instances of the GADT than
  ways to construct its other type instances.)  Further, to ensure
  that the categorical Lans interpreting $\Lan$-constructs are
  $\omega$-cocontinuous we also require that $K$ is a polynomial
  functor in the $\alpha$s that does not involve any constant
  functors. This means that our constructors have types of the form
  $F\ol\beta \to F(K\ol\beta)$ with no constants in $K$, so that
  exactly the corresponding instances of the GADT are inhabited for
  every instantiation of the $\alpha$s, and the objects for each
  instantiation of the $\alpha$s are constructed in exactly the same
  way. That is, the inhabited instances and the construction of the
  data in those instances are type independent.}
